---
title: "ESM 232 Assignment 3"
format: html
author: "Ella Moore and Sneha Kumar"
---

Step 1: Read the climate data from the txt file into a data table 
```{r}
library(here)
library(dplyr)
clim_data<-read.table(here("Assignments/clim.txt"), header = T)
```

Step 2: Clean the climate data into two new data frames 
The first includes only the average minimum temperature (C) for the month of February 
The second contains the average precipitation (mm) for the month of January
```{r}
#Calculate the annual average minimum temp in Feb 
feb_min_temp <- clim_data |>
  filter(month == "2") |>
  select(day, month, year, tmin_c)
feb_min_temp

avg_min_temp <- feb_min_temp |>
  group_by(year) |>
  summarize(avg_temp = mean(tmin_c, na.rm = TRUE))
avg_min_temp

#Calculate the annual average precipitation in Jan 
jan_precipitation <- clim_data |>
  filter(month == "1") |>
  select(day, month, year, precip)
jan_precipitation

precipsum <- jan_precipitation |>
  group_by(year) |>
  summarize(sumprecip = sum(precip, na.rm = TRUE))
precipsum
```
Step 3: Call the function and run it to fill the data frame to identify annual yield anomalies
```{r}
# Call the function located in the R folder
source(here("R/HW3.R"))

# Merge the data frames into one and create new blank column
inputsandoutputs <- merge(avg_min_temp, precipsum, by="year") |>
  mutate(yieldvalues = NA)

yields <- yield_anomaly(inputsandoutputs$avg_temp, inputsandoutputs$sumprecip)

inputsandoutputs$yieldvalues <- yields$yield

inputsandoutputs
```

Step 4: Create a summary table to show the mean, minimum, and maximum yield anomaly values
```{r}
summarytable <- data.frame(
  StatisticType = c("Mean Yield Anomaly", "Min Yield Anomaly", "Max Yield Anomaly"),
  Values = c(yields$mean, yields$min, yields$max)
)
print(summarytable)
```

